Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-15 chapters/LBM_technique.tex
addressed notation and made font size larger in stress-time curve

The lattice boltzmann method works by evolving a population distribution $f_{i}(\mathbf{x}, t)$ on a cubic lattice with 
timestep $\Delta t$ with lengthscale $\Delta x$. \cite{qian_lattice_1992, succi_lattice_2018, he_theory_1997} The cubic lattice or stencil is described
DnQm where n and m represent the number of dimensions and directions in the stencil. \cite{succi_lattice_2018, schmieschek_lb3d_2017}
Common stencils include the D1Q5, D2Q9 and D3Q19 stencil, all of which recover mass and momentum conservation and are Galilean invariant.
For energy conservation, a higher order stencil such as D3Q27 is necessary. The D2Q9 and D3Q19 stencils have 9 and 19 populations respectively that include 
connections to nearest and next nearest neighbour points. The D3Q19 
velocity set is used in this work with the indexes $i$ represent each of the 19 velocities and conserves mass and momentum 
to the second order. It can be seen in Figure \ref{fig:d3q19_lattice}.

@@

\begin{equation}
    f_{i}(\mathbf{x} + \mathbf{c}_{i}\Delta t, t + \Delta t) = f_{i}(\mathbf{x}, t) - \frac{1}{\tau}(f_{i}(\mathbf{x}, t) 
    - f_{i}^{eq}(\mathbf{x}, t))
    \label{eq:LBM_BGK}
\end{equation}

@@

$\rho$ and $\textbf{u}$ in Equation \ref{eq:LBM_Feq} are defined as the macroscopic parameters for density and velocity 
and can be calculated from the mass distribution using $\rho = \sum f_i$ and $\rho \mathbf{u} = \sum f_i \mathbf{c}_i$ 
respectively. Using the Chapman-Enskog expansion, the Navier-Stokes equation at the incompressible limit at low Mach
numbers can be recovered. \cite{qian_lattice_1992, he_lattice_1997} The regime accessible through this method is suitable 
for simulations in this work as the $ 1 \leq Re \leq 100 $ and the mach number, $Ma < 0.01$, fulfilling the stability 
criterion and usage requirements for the presented hydrodynamic model.

@@

\begin{equation}
    \begin{split}
        \rho &= \sum_i f_i \\
        \rho\vec{u} &= \sum_i f_i \mathbf{c_i}
    \end{split}
\end{equation}

@@

\begin{equation}
    \begin{split}
        % \Pi^{(0)}_{\alpha \beta} &= \sum_i \mathbf{c_i}_\alpha \mathbf{c_i}_\beta f_i^{eq} = p \delta_{\alpha \beta} + \rho \vec{u}_\alpha \vec{u}_\beta \\
        % \Pi^{(1)}_{\alpha \beta} &= (1 - \frac{1}{2 \tau})\sum_i \mathbf{c_i}_\alpha \mathbf{c_i}_\beta f_i^{eq} = \nu (\nabla_{\alpha}(\rho \vec{\beta}) + \nabla_{\beta}(\rho \vec{\alpha}))
        \Pi^{(0)} &= \sum_i f_i^{eq} \vec{c}_i \otimes \vec{c}_i = p\mathbb{I} + \rho \vec{u} \otimes \vec{u} \\
        \Pi^{(1)} &= (1 - \frac{1}{2 \tau})\sum_i f_i^{neq} \vec{c}_i \otimes \vec{c}_i = -\eta (\nabla \vec{u} + (\nabla \vec{u})^T)
    \end{split}
\end{equation}

@@

The SC model mimics Cahn-Hilliard type behaviour through a force applied from the other fluid species $k'$ in adjacent 
cells $\mathbf{x'}$ on fluid species $k$ at point $\mathbf{x}$. \cite{shan_lattice_1993, shan_simulation_1994, 
shan_multicomponent_1995, he_discrete_1998, jansen_bijels_2011, chin_lattice_2002} Both fluid species are defined
by their own distribution equation defined in Equation \ref{eq:LBM_BGK} The strength of this force is controlled 
through an interaction parameter, $g_{kk'}$ with no contribution from self interaction of the fluid as these are 
set to zero. The SC force can then be written out in Equation \ref{eq:sc_model}.

\begin{equation}
% F_{k}^{SC}(\mathbf{x}, t) = -\Psi^{k}(\mathbf{x}, t)\sum_{k'}g_{cc'}\sum_{\mathbf{x'}}\Psi_{k'}(\mathbf{x'}, t)(\mathbf{x'} 
% - \mathbf{x})
\vec{F}_k(\vec{x}) \Delta t = - \sum_{k'} \sum_i \frac{w_i}{c_s^2} g_{kk'} \psi_k(\vec{x})\psi_{k'}(\vec{x}+\vec{c}_i) \vec{c}_i
\label{eq:sc_model}
\end{equation}

An effective mass of each fluid at node $\mathbf{x}$ is used in place of the actual density to scale it between zero 
and one and is defined as $\psi^{k}(\mathbf{x},t) = \rho_{0}\left[1 - \exp(-\frac{\rho^{k}(\mathbf{x}, t)}{\rho_{0}})\right]$. 
In this model, the SC force is incorporated into the macroscopic velocities that are then used to calculate the equilibrium
distribution $f_{i}^{k, eq}$ for fluid $k$, defined as,

@@

\begin{equation}
    \begin{split}
    \vec{F_p} = m_p \frac{\vec{u}_p}{dt} , \\
    \vec{D_p} = \mathbf{J}_p \frac{\vec{\omega}_p}{dt} ,
    \label{eq:md}
    \end{split}
\end{equation}

$\mathbf{F_p}$ and $\mathbf{D_p}$ represent the force and torque acting on a particle with mass $m_p$ and moment of inertia 
$\mathbf{J}_p$. $\mathbf{u}_p$ and $\mathbf{\omega_{p}}$ are the linear and angular velocities of the particle. The equations of 
motion are evolved over time using a leapfrog integrator. \cite{jansen_bijels_2011}

@@

where $R_i$ and $R_j$ are the radii of each particle involved in the interaction, $\vec{r}_{ij}$ is the distance
vector between the particle centers, $\mathbf{u}_{ij}$ are the relative velocities of the particles and $\Delta_c$ 
is the cutoff distance when the lubrication force begins to act. If particles are able to overcome the lubrication forces, 
a hertzian contact force is also added to ensure that there is no particle overlap, defined in Equation \eqref{eq:hertz}

\begin{equation}
    \phi_{H} = K_{H}(R_i + R_j - |\mathbf{r}_{ij}|)^{5/2}, r < R_i + R_j
    \label{eq:hertz}
\end{equation}

@@

Where $\mu_0 = 4\pi \cdot 10^{-7} \frac{H}{m}$,  $\Hat{\mathbf{o_i}}$ is the orientation unit vector of particle 
$i$, $\Hat{\vec{r}_{ij}}$ is the distance vector between particles $i$ and $j$ and $m_i$ is the magnitude the 
magnetic dipole of particle $i$. From the potential, the force and torque of the dipole force between particles 
can be found. These expressions are shown in equations \ref{eq:dipole_magnetic_force} and \ref{eq:dipole_magnetic_torque} 
for the force and torque respectively.

@@

\begin{table}[h!]
\centering
\begin{tabular}{||c c c c c c c c c c c c c||} 
 \hline
 $t_{sim}$ & $L_x$ & $\rho_f $ & $\tau$ & $\eta_f $ & $g_{cc'} \Delta m(\Delta t)^2/(\Delta x)^5$ & $\sigma$ & $\phi_f$ & $\rho_p$ & $\phi_p$ & $V_p (\Delta x)^3$ & $d_c \Delta x$ & $K_H$ \\ [0.5ex] 
 \hline\hline
 $10^5 \Delta t$ & 256$\Delta x$ & 0.7 $\frac{\Delta m}{(\Delta x)^3}$ & 1$\Delta t$ & 0.117 $\frac{\Delta m}{\Delta x \Delta t}$ & 0.08 & 0.0267 $\frac{\Delta m}{(\Delta t)^2}$ & 0.5 & 1 $\frac{\Delta m}{(\Delta x)^3}$ & 0.15 & 2000$\pi (\Delta x)^3$ & 2/3 $\Delta x$& 100 $(\Delta x)^{1/2}(\Delta t)^2/\Delta m$\\ [1ex] 
 \hline
\end{tabular}
\caption{Summary of the fluid, particle and model parameters used in the proposed simulations}
\label{table:model_params}
\end{table}

In these simulations, the aspect ratio is defined as $\alpha = \frac{R_{\parallel}}{R_{\perp}}$ with $\alpha = 0.5, 1, 2$ 
used in these simulations. All particle geometries are volume matched, based on the spherical particle volume. For the 
spherical particle, the radius was set at $R_{\parallel} = R_{\perp} = 7.9$ which sets the $\alpha = 0.5$ particles to 
$R_{\parallel} = 5, R_{\perp} = 10$ and the $\alpha = 2$ particles to $R_{\parallel} = 12.6, R_{\perp} = 6.3$. 
$R_{\parallel}$ and $R_{\perp}$ is the radius of the particle parallel and perpendicular to the symmetry axis of 
the particle respectively. These values are identical to what was used in Gunther et al, allowing comparisons to the 
trends they observed. \cite{gunther_timescales_2014} The particles were randomly placed in the system and an equilibration
step was performed to push the particles further away from each after placement. Once equilibration is performed, a 
constant magnetic field is switched on with dimensionless strengths defined using $\bar{B}$.

@@

Bijel templates have been synthesized using water/2-6-lutidine for further applications. \cite{lee_making_2013} 
While intrinsically polymerizable bijels have been prepared using oligomeric polymers, STRIPS and VIPS currently 
utilize small molecule liquids in their bijel casting mixtures. Therefore, an experimentally viable system using 
water/2-6-lutidine will be discussed in the context of the work presented here. The critical temperature (UCST) 
for this system is $T_c = 34 ^{\circ}C$, with many systems heated to $T = 40^{\circ}C$ to form the bijel. At this 
temperature, the fluid properties are $\sigma = 0.22 mN/m$ and $\eta = 2.38 mPas$. \cite{grattoni_lower_1993} 
Fei et al. developed a method to synthesize polystyrene microparticles ($(R_p = 2 \mu m)$) with tunable geometries 
by mechanically stretching or compressing them, then coating them with nickel to create magnetic microparticles with 
a dipole moment of approximately$\approx 3 \cdot 10^{-14} Am^2$. \cite{fei_active_2017, fei_magneto-capillary_2020} 
Comparing the physical properties to simulation parameters, the simulations correspond to a length scale of
$\Delta x = 252 \cdot 10^{-9} \text{m}$, timescale of $\Delta t = 624 \cdot 10^{-9} \text{s}$, 
mass scale of $\Delta m = 2.24 \cdot 10^{-15} \text{kg}$ and charge scale of $\Delta q = 2.948 \cdot 10^{-7} \text{C}$.

@@

Using the fluid and non-ideal mixing parameters defined in Table \ref{table:model_params}, a single water in oil 
droplet, was initialized in the center of the system and run till the droplet radii reached steady state. 
This process was repeated for various droplet diameters, specified between $70$ and $85 \%$ of the system size at $5\%$ 
increments. The system had a box length of $L = 64$ and was run for $50000$ timesteps with periodic boundary conditions 
on all sides. Data was dumped every 10000 timesteps with only the last timestep used to calculate the values shown. Due 
to the diffuse interface present in this method, the density profiles were verified to be at steady state at the end of 
the simulation before further analysis was performed. \cite{frijters_effects_2012} Past work also showed that the 
pressures remained constant after 5 lattice units away from the interface. \cite{frijters_effects_2012}

The pressure difference was calculated using the difference in scalar pressure between the center of the droplet and 
one of the corners of the box. Two techniques exist to calculate the radius of the droplet. The first directly measures 
the distance between opposite points of the droplet as measured from the density profile at opposite points of the droplet, 
and the second is based upon the total mass and density of the fluid making up the droplet. The second method offers greater 
accuracy by avoiding discretization errors that may exist in the first method, and also allows for the correction of 
particle volume at the interface if the system contained them. First, a local effective mass density 
$\rho^b_{eff} = \rho^{b}_{d} - \rho^{b}_{m}$ is found by calculating the difference in density between the blue 
fluid in the droplet and in the matrix. Next, the droplet mass is calculated as $M_d = \sum_{\mathbf{x}}{\rho_{eff}^{b}}$. 
The radius of the droplet can then be calculated from the mass and density of the droplets as $R_d = \sqrt[3]{\frac{3}{4\pi} 
\frac{M_d}{\rho^b_d - \rho^b_m}}$. To linearize the plot, the $\Delta P$ was plotted against $\frac{2}{R_d}$ with 
$\sigma$ being the slope of the fit.

>>>>

The lattice boltzmann method works by evolving a population distribution $f_{i}(\vec{x}, t)$ on a cubic lattice with 
timestep $\Delta t$ with lengthscale $\Delta x$. \cite{qian_lattice_1992, succi_lattice_2018, he_theory_1997} The cubic lattice or stencil is described
DnQm where n and m represent the number of dimensions and directions in the stencil. \cite{succi_lattice_2018, schmieschek_lb3d_2017}
Common stencils include the D1Q5, D2Q9 and D3Q19 stencil, all of which recover mass and momentum conservation and are Galilean invariant.
For energy conservation, a higher order stencil such as D3Q27 is necessary. The D2Q9 and D3Q19 stencils have 9 and 19 populations respectively that include 
connections to nearest and next nearest neighbour points. The D3Q19 
velocity set is used in this work with the indexes $i$ represent each of the 19 velocities and conserves mass and momentum 
to the second order. It can be seen in Figure \ref{fig:d3q19_lattice}.

@@

\begin{equation}
    f_{i}(\vec{x} + \vec{c}_{i}\Delta t, t + \Delta t) = f_{i}(\vec{x}, t) - \frac{1}{\tau}(f_{i}(\vec{x}, t) 
    - f_{i}^{eq}(\rho, \vec{u}))
    \label{eq:LBM_BGK}
\end{equation}

@@

$\rho$ and $\textbf{u}$ in Equation \ref{eq:LBM_Feq} are defined as the macroscopic parameters for density and velocity 
and can be calculated from the mass distribution using $\rho = \sum f_i$ and $\rho \vec{u} = \sum f_i \vec{c}_i$ 
respectively. Using the Chapman-Enskog expansion, the Navier-Stokes equation at the incompressible limit at low Mach
numbers can be recovered. \cite{qian_lattice_1992, he_lattice_1997} The regime accessible through this method is suitable 
for simulations in this work as the $ 1 \leq Re \leq 100 $ and the mach number, $Ma < 0.01$, fulfilling the stability 
criterion and usage requirements for the presented hydrodynamic model.

@@

\begin{equation}
    \begin{split}
        \rho &= \sum_i f_i \\
        \rho\vec{u} &= \sum_i f_i \vec{c}_i
    \end{split}
\end{equation}

@@

\begin{equation}
    \begin{split}
        \Pi^{(0)} &= \sum_i f_i^{eq} \vec{c}_i \otimes \vec{c}_i = p\mathbb{I} + \rho \vec{u} \otimes \vec{u} \\
        \Pi^{(1)} &= (1 - \frac{1}{2 \tau})\sum_i f_i^{neq} \vec{c}_i \otimes \vec{c}_i = -\eta (\nabla \vec{u} + (\nabla \vec{u})^T)
    \end{split}
\end{equation}

@@

The SC model mimics Cahn-Hilliard type behaviour through a force applied from the other fluid species $k'$ in adjacent 
cells $\vec{x'}$ on fluid species $k$ at point $\vec{x}$. \cite{shan_lattice_1993, shan_simulation_1994, 
shan_multicomponent_1995, he_discrete_1998, jansen_bijels_2011, chin_lattice_2002} Both fluid species are defined
by their own distribution equation defined in Equation \ref{eq:LBM_BGK} The strength of this force is controlled 
through an interaction parameter, $g_{kk'}$ with no contribution from self interaction of the fluid as these are 
set to zero. The SC force can then be written out in Equation \ref{eq:sc_model}.

\begin{equation}
\vec{F}_k(\vec{x}) \Delta t = - \sum_{k'} \sum_i \frac{w_i}{c_s^2} g_{kk'} \psi_k(\vec{x})\psi_{k'}(\vec{x}+\vec{c}_i) \vec{c}_i
\label{eq:sc_model}
\end{equation}

An effective mass of each fluid at node $\vec{x}$ is used in place of the actual density to scale it between zero 
and one and is defined as $\psi^{k}(\vec{x},t) = \rho_{0}\left[1 - \exp(-\frac{\rho^{k}(\vec{x}, t)}{\rho_{0}})\right]$. 
In this model, the SC force is incorporated into the macroscopic velocities that are then used to calculate the equilibrium
distribution $f_{i}^{k, eq}$ for fluid $k$, defined as,

@@

\begin{equation}
    \begin{split}
    \vec{F_p} = m_p \frac{\vec{u}_p}{dt} , \\
    \vec{D_p} = \vec{J}_p \frac{\vec{\omega}_p}{dt} ,
    \label{eq:md}
    \end{split}
\end{equation}

$\vec{F_p}$ and $\vec{D_p}$ represent the force and torque acting on a particle with mass $m_p$ and moment of inertia 
$\vec{J}_p$. $\vec{u}_p$ and $\vec{\omega_{p}}$ are the linear and angular velocities of the particle. The equations of 
motion are evolved over time using a leapfrog integrator. \cite{jansen_bijels_2011}

@@

where $R_i$ and $R_j$ are the radii of each particle involved in the interaction, $\vec{r}_{ij}$ is the distance
vector between the particle centers, $\vec{u}_{ij}$ are the relative velocities of the particles and $\Delta_c$ 
is the cutoff distance when the lubrication force begins to act. If particles are able to overcome the lubrication forces, 
a hertzian contact force is also added to ensure that there is no particle overlap, defined in Equation \eqref{eq:hertz}

\begin{equation}
    \phi_{H} = K_{H}(R_i + R_j - |\vec{r}_{ij}|)^{5/2}, r < R_i + R_j
    \label{eq:hertz}
\end{equation}

@@

Where $\mu_0 = 4\pi \cdot 10^{-7} \frac{H}{m}$,  $\Hat{\vec{o_i}}$ is the orientation unit vector of particle 
$i$, $\Hat{\vec{r}_{ij}}$ is the distance vector between particles $i$ and $j$ and $m_i$ is the magnitude the 
magnetic dipole of particle $i$. From the potential, the force and torque of the dipole force between particles 
can be found. These expressions are shown in equations \ref{eq:dipole_magnetic_force} and \ref{eq:dipole_magnetic_torque} 
for the force and torque respectively.

@@

% \begin{table}[h!]
% \centering
% \begin{tabular}{||c c c c c c c c c c c c c||} 
%  \hline
%  $t_{sim}$ & $L_x$ & $\rho_f $ & $\tau$ & $\eta_f $ & $g_{cc'} \Delta m(\Delta t)^2/(\Delta x)^5$ & $\sigma$ & $\phi_f$ & $\rho_p$ & $\phi_p$ & $V_p (\Delta x)^3$ & $d_c \Delta x$ & $K_H$ \\ [0.5ex] 
%  \hline\hline
%  $10^5 \Delta t$ & 256$\Delta x$ & 0.7 $\frac{\Delta m}{(\Delta x)^3}$ & 1$\Delta t$ & 0.117 $\frac{\Delta m}{\Delta x \Delta t}$ & 0.08 & 0.0267 $\frac{\Delta m}{(\Delta t)^2}$ & 0.5 & 1 $\frac{\Delta m}{(\Delta x)^3}$ & 0.15 & 2000$\pi (\Delta x)^3$ & 2/3 $\Delta x$& 100 $(\Delta x)^{1/2}(\Delta t)^2/\Delta m$\\ [1ex] 
%  \hline
% \end{tabular}
% \caption{Summary of the fluid, particle and model parameters used in the proposed simulations}
% \label{table:model_params}
% \end{table}

\begin{table}
    \centering
    \caption{Summary of the parameters corresponding to the fluid, particle and model used in the simulations}
    \label{table:model_params}
    \begin{tabular}{|c|r|}
    \hline
    % Parameter & Value \\
    % \hline
    $L_x$ & 256 $\Delta x$ \\
    $\rho_f$ & 0.7 $ \frac{\Delta m}{(\Delta x)^3}$\\
    $\nu_f$ & 1/6 $\frac{\Delta x}{(\Delta t)^2}$ \\
    $\sigma $ & 0.0267 $\frac{\Delta m}{(\Delta t)^2}$ \\
    $\tau$ & 1 $\Delta t$\\
    $g_{kk'}$ & 0.08 $\frac{(\Delta x)^5}{\Delta m(\Delta t)^2}$\\
    $\phi_f$ & 0.5 \\
    $\phi_p$ & 0.15 \\
    $n_p$ & 1200 \\
    $\rho_p$ & 1 $ \frac{\Delta m}{(\Delta x)^3}$\\
    $V_p$ & $2000\pi/3 (\Delta x)^3$ \\
    $m$ & 1 $\frac{\Delta q}{\Delta t(\Delta x)^2}$\\
    $d_c$ & 2/3 $\Delta x$ \\
    $K_H$ & 100 $\frac{\Delta m}{(\Delta x)^{1/2}(\Delta t)^2} $\\
    \hline
    \end{tabular}
\end{table}

In these simulations, the aspect ratio is defined as $\alpha = \frac{R_{\parallel}}{R_{\perp}}$ with $\alpha = 0.5, 1, 2$ 
used in these simulations. All particle geometries are volume matched, based on the spherical particle volume. For the 
spherical particle, the radius was set at $R_{\parallel} = R_{\perp} = 7.9 \Delta x$ which sets the $\alpha = 0.5$ particles to 
$R_{\parallel} = 5 \Delta x, R_{\perp} = 10 \Delta x$ and the $\alpha = 2$ particles to $R_{\parallel} = 12.6, R_{\perp} = 6.3 \Delta x$. 
$R_{\parallel}$ and $R_{\perp}$ is the radius of the particle parallel and perpendicular to the symmetry axis of 
the particle respectively. These values are identical to what was used in Gunther et al, allowing comparisons to the 
trends they observed. \cite{gunther_timescales_2014} The particles were randomly placed in the system and an equilibration
step was performed to push the particles further away from each after placement. Once equilibration is performed, a 
constant magnetic field is switched on with dimensionless strengths defined using $\bar{B}$.

@@

Bijel templates have been synthesized using water/2-6-lutidine for further applications. \cite{lee_making_2013} 
While intrinsically polymerizable bijels have been prepared using oligomeric polymers, STRIPS and VIPS currently 
utilize small molecule liquids in their bijel casting mixtures. Therefore, an experimentally viable system using 
water/2-6-lutidine will be discussed in the context of the work presented here. The critical temperature (UCST) 
for this system is $T_c = 34 ^{\circ}\text{C}$, with many systems heated to $T = 40^{\circ}\text{C}$ to form the bijel. At this 
temperature, the fluid properties are $\sigma = 0.22 \text{mN/m}$ and $\eta = 2.38 \text{mPas}$. \cite{grattoni_lower_1993} 
Fei et al. developed a method to synthesize polystyrene microparticles ($R_p = 2\,\mu\text{m}$) with tunable geometries 
by mechanically stretching or compressing them, then coating them with nickel to create magnetic microparticles with 
a dipole moment of approximately$\approx 3 \cdot 10^{-14} \text{Am}^2$. \cite{fei_active_2017, fei_magneto-capillary_2020} 
Comparing the physical properties to simulation parameters, the simulations correspond to a length scale of
$\Delta x = 252 \cdot 10^{-9} \text{m}$, timescale of $\Delta t = 624 \cdot 10^{-9} \text{s}$, 
mass scale of $\Delta m = 2.24 \cdot 10^{-15} \text{kg}$ and charge scale of $\Delta q = 2.948 \cdot 10^{-7} \text{C}$.

@@

Using the fluid and non-ideal mixing parameters defined in Table \ref{table:model_params}, a single water in oil 
droplet, was initialized in the center of the system and run till the droplet radii reached steady state. 
This process was repeated for various droplet diameters, specified between 70 and 85 \% of the system size at 5\% 
increments. The system had a box length of $L = 64 \Delta x$ and was run for 50000 timesteps with periodic boundary conditions 
on all sides. Data was dumped every 10000 timesteps with only the last timestep used to calculate the values shown. Due 
to the diffuse interface present in this method, the density profiles were verified to be at steady state at the end of 
the simulation before further analysis was performed. \cite{frijters_effects_2012} Past work also showed that the 
pressures remained constant after 5 lattice units away from the interface. \cite{frijters_effects_2012}

The pressure difference was calculated using the difference in scalar pressure between the center of the droplet and 
one of the corners of the box. Two techniques exist to calculate the radius of the droplet. The first directly measures 
the distance between opposite points of the droplet as measured from the density profile at opposite points of the droplet, 
and the second is based upon the total mass and density of the fluid making up the droplet. The second method offers greater 
accuracy by avoiding discretization errors that may exist in the first method, and also allows for the correction of 
particle volume at the interface if the system contained them. First, a local effective mass density 
$\rho^b_{eff} = \rho^{b}_{d} - \rho^{b}_{m}$ is found by calculating the difference in density between the blue 
fluid in the droplet and in the matrix. Next, the droplet mass is calculated as $M_d = \sum_{\vec{x}}{\rho_{eff}^{b}}$. 
The radius of the droplet can then be calculated from the mass and density of the droplets as $R_d = \sqrt[3]{\frac{3}{4\pi} 
\frac{M_d}{\rho^b_d - \rho^b_m}}$. To linearize the plot, the $\Delta P$ was plotted against $\frac{2}{R_d}$ with 
$\sigma$ being the slope of the fit.

### 2025-04-15 chapters/paper1_new.tex
Making notation consistent
addressed notation and made font size larger in stress-time curve

\section[Introduction]{Introduction\protect\footnote{This section is based on work published in the Royal Society of Chemistry's Soft Matter journal. It is reproduced from Ref.~\cite{karthikeyan_formation_2024} with permission from the Royal Society of Chemistry.}}

Emulsions formed from mixtures of immiscible liquids are widely used as structural 
templates for fabricating porous materials. These include a broad array of applications 
such as porous polymers, ceramics, synthetic nanoparticles, and engineered tissue scaffolds 
\cite{zhang_emulsion_2019, stubenrauch_emulsion_2018, jones_high-temperature_2009, binks_macroporous_2002, aldemir_dikici_basic_2020}. 
The technique of emulsion templating 
provides a flexible and efficient approach to controlling key structural features of the final
material—namely, the domain size, volume fraction, and phase composition—by manipulating the 
formulation and processing conditions of the emulsion \cite{mudassir_fundamentals_2021}. To ensure 
thermodynamic and mechanical stability during processing, emulsions are typically stabilized by additives. 
Among these, particle-stabilized systems—commonly referred to as Pickering emulsions—have emerged as attractive 
alternatives to traditional surfactant-based emulsions due to their ability to form stable interfaces without 
relying on potentially harmful chemicals \cite{ramsden_separation_1904, zheng_pickering_2022, menner_particle-stabilized_2007}.

@@

Beyond chemical composition, the geometry of the stabilizing particles plays a critical role in determining 
the microstructure and stability of emulsions. When anisotropic particles—such as ellipsoids or rods—adsorb at 
fluid interfaces, they give rise to shape-dependent capillary interactions that are markedly different from those 
induced by spherical particles \cite{loudet_capillary_2005,madivala_exploiting_2009}. These interactions can 
significantly enhance emulsion stability, as demonstrated by Madivala et al., and often result in distinct particle 
arrangements at the interface \cite{madivala_self-assembly_2009}. Moreover, anisotropic particles are capable of 
self-assembling in response to interface curvature gradients, where the curvature acts as an external field influencing 
particle orientation and spatial distribution \cite{cavallaro_curvature-driven_2011, furst_directing_2011}. Computational 
studies, including Monte Carlo and Lattice Boltzmann simulations, have provided insight into how such particles dynamically 
interact with fluid interfaces \cite{bresme_orientational_2007, bresme_computer_2008, gunther_lattice_2013,gunther_timescales_2014}. 
These simulations show that anisotropic particles introduce additional timescales during 
emulsion formation due to their rotational behavior. Initially, they rotate to align their major axes with the interface, 
and later reorient under capillary forces, contributing to prolonged coarsening and finer domain features. Experimental studies 
by Hijnen et al.~\cite{hijnen_bijels_2015} confirmed that rod-shaped particles can jam more quickly than spheres at smaller 
interfacial separations and can even tilt out of the plane due to compression during domain evolution.

>>>>

\section[Introduction]{Introduction\protect\footnote{This chapter is based on work published in the Royal Society of Chemistry's Soft Matter journal. It is reproduced from Ref.~\cite{karthikeyan_formation_2024} with permission from the Royal Society of Chemistry.}}

Emulsions formed from mixtures of immiscible liquids are widely used as structural 
templates for fabricating porous materials. These include a broad array of applications 
such as porous polymers, ceramics, synthetic nanoparticles, and engineered tissue scaffolds 
\cite{zhang_emulsion_2019, stubenrauch_emulsion_2018, jones_high-temperature_2009, binks_macroporous_2002, aldemir_dikici_basic_2020}. 
The technique of emulsion templating 
provides a flexible and efficient approach to controlling key structural features of the final
material such as the domain size, volume fraction, and phase composition—by manipulating the 
formulation and processing conditions of the emulsion \cite{mudassir_fundamentals_2021}. To ensure 
thermodynamic and mechanical stability during processing, emulsions are typically stabilized by additives. 
Among these, particle-stabilized systems or Pickering emulsions have emerged as attractive 
alternatives to traditional surfactant-based emulsions due to their ability to form stable interfaces without 
relying on potentially harmful chemicals \cite{ramsden_separation_1904, zheng_pickering_2022, menner_particle-stabilized_2007}.

@@

Beyond chemical composition, the geometry of the stabilizing particles plays a critical role in determining 
the microstructure and stability of emulsions. When anisotropic particles such as ellipsoids adsorb at 
fluid interfaces, they give rise to shape-dependent capillary interactions that are markedly different from those 
induced by spherical particles \cite{loudet_capillary_2005,madivala_exploiting_2009}. These interactions can 
significantly enhance emulsion stability, as demonstrated by Madivala et al., and often result in distinct particle 
arrangements at the interface \cite{madivala_self-assembly_2009}. Moreover, anisotropic particles are capable of 
self-assembling in response to interface curvature gradients, where the curvature acts as an external field influencing 
particle orientation and spatial distribution \cite{cavallaro_curvature-driven_2011, furst_directing_2011}. Computational 
studies, including Monte Carlo and Lattice Boltzmann simulations, have provided insight into how such particles dynamically 
interact with fluid interfaces \cite{bresme_orientational_2007, bresme_computer_2008, gunther_lattice_2013,gunther_timescales_2014}. 
These simulations show that anisotropic particles introduce additional timescales during 
emulsion formation due to their rotational behavior. Initially, they rotate to align their major axes with the interface, 
and later reorient under capillary forces, contributing to prolonged coarsening and finer domain features. Experimental studies 
by Hijnen et al.~\cite{hijnen_bijels_2015} confirmed that rod-shaped particles can jam more quickly than spheres at smaller 
interfacial separations and can even tilt out of the plane due to compression during domain evolution.

### 2025-04-15 chapters/paper3.tex
addressed notation and made font size larger in stress-time curve

The rheological properties of bijels are critical to their processability and structural fidelity across a variety of manufacturing techniques. In solvent transfer-induced phase separation 
(STrIPS), for example, the flow rate of the casting mixture and the co-solvent ratio play a key role in controlling the microstructure of the bijel and determining whether droplets or 
rope-like structures emerge during synthesis \cite{haase_continuous_2015, haase_situ_2016}. Similarly, in volatile-induced phase separation (VIPS), the viscosity of the solvent rich and poor 
phases determines the extent of phase separation before gelation or jamming occurs, ultimately governing final domain morphology \cite{wang_scalable_2020}. In homogenization-driven methods, 
the imposed shear rate directly influences the interfacial jamming dynamics and can either assist or disrupt bicontinuous structure formation depending on processing windows 
\cite{huang_bicontinuous_2017}. Understanding these rheological dependencies is essential for tuning bijel architectures for advanced applications.

>>>>

The rheological properties of bijels are critical to their processability and structural fidelity across a variety of manufacturing techniques. In solvent transfer-induced phase separation 
(STrIPS), for example, the flow rate of the casting mixture and the co-solvent ratio play a key role in controlling the microstructure of the bijel and determining whether droplets or 
rope-like structures emerge during synthesis \cite{haase_continuous_2015, haase_situ_2016}. Similarly, in vapor-induced phase separation (VIPS), the viscosity of the solvent rich and poor 
phases determines the extent of phase separation before gelation or jamming occurs, ultimately governing final domain morphology \cite{wang_scalable_2020}. In homogenization-driven methods, 
the imposed shear rate directly influences the interfacial jamming dynamics and can either assist or disrupt bicontinuous structure formation depending on processing windows 
\cite{huang_bicontinuous_2017}. Understanding these rheological dependencies is essential for tuning bijel architectures for advanced applications.
